Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-05-11 paper/manuscript.Rmd
P5

Forage limitation and over-grazing is a major concern for caribou populations:
it can reduce female reproductive success [@Schaefer2016] and thus could lead to
movement between islands. Forage scarcity has been proposed as a potential
reason for caribou moving between arctic islands on the sea-ice in winter
[@Miller1977]. Caribou were introduced to Fogo Island (n = 26 animals between
1964–67) and Change Island (n = 5 animals in 1964) from Newfoundland as part of
a series of translocations and introductions throughout the province (Bergerud &
Mercer 1989). It is unknown when caribou began colonizing nearby islands, but
given that some are small, over-grazing by newly arrived caribou may rapidly
deplete forage on an annual cycle [@Bergerud1990]. Although unconfirmed, it is
also possible that Fogo Island has experienced forage depletion since the
introduction of caribou in the 1960s. More likely, however, is that fine-scale
competition for foraging resources drives caribou to swim to new islands.

>>>>

Caribou were introduced to Fogo Island (n = 26 animals between 1964–67) and
Change Island (n = 5 animals in 1964) from Newfoundland as part of a series of
translocations and introductions throughout the province [@Bergerud1989]. Typical caribou predators such as black bears (*Ursus americanus*) and
wolves (*Canis lupus*) are absent from Fogo Island; the dominant predator is
coyote (*Canis latrans*), which predate calves but it remains unclear what threat
they pose, if any, to adult caribou. Given the minimal risk of predation to adults,
we propose that forage limitation is more likely to drive movement between
islands than predator avoidance. For several decades after their introduction to
Fogo Island, the caribou population did not exceed ~100 individuals (Bergerud
and Mercer 1989, Newfoundland and Labrador Wildlife Division, unpublished data).
During the 1990s, population density reached ~300 animals; anecdotal evidence
suggests caribou were first seen on nearby islands during this same period.
Increased competition for resources on Fogo due to this increase in density
could have precipitated movement to other islands.

Forage limitation and over-grazing is a major concern for caribou populations:
it can reduce female reproductive success [@Schaefer2016] and thus could lead to
movement between islands. Forage scarcity has been proposed as a potential
reason for caribou moving between arctic islands on the sea-ice in winter
[@Miller1977].

It is unknown when caribou began colonizing nearby islands, but
given that some are small, over-grazing by newly arrived caribou may rapidly
deplete forage on an annual cycle [@Bergerud1990]. Although unconfirmed, it is
also possible that Fogo Island has experienced forage depletion since the
introduction of caribou in the 1960s. More likely, however, is that fine-scale
competition for foraging resources drives caribou to swim to new islands.
